---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rbounds: Estimating Bounds on Partially Identified Parameters

The goal of rbounds is to estimate bounds on the location of parameters
in models in which the value of the parameter can only be partially
identified.

The package implements non-parametric bounds estimation following the
work of Charles Manski on partial identification. It uses kernel
regression estimators from the `np` package to estimate the
identifiable components of the bounds.

**Current functionality:**

1. `pidoutcomes()`: Bounds on conditional means with missing outcomes
2. `ate_bounds()`: Manski bounds on Average Treatment Effects (ATE)

## Example of bounding the conditional mean

Imagine that we are interested in estimating
$E(y | x)$ where $y \in \{0,1\}$. Unfortunately, we are 
missing some observations of $y$. Whether an observation
is missing is denoted by a dummy variable $z$, where when $z=1$
$y$ is observed and otherwise the value of $y$ is not observed.
We have some covariates $x$ whose value we observe for all observations.

By the law of total probability it follows that 

\[E(y | x) = E(y | x, z = 1)P(z = 1 | x) + E(y | x, z = 0)P(z = 0 | x).\]

Each of the pieces of this equation can be estimated from 
the available data with the exception of $E(y|x, z=0)$, which
is the conditional expectation of the outcome for the cases in 
which the outcome value is unobserved. 

However, since $0 \leq y \leq 1$, the unobserved
function can take on the value of at most $1$ -- if all the
unobserved outcomes were 1, or at least $0$. Thus the true
conditional mean must fall within the bounds:

\[
E(y | x, z = 1)P(z = 1 | x) \leq E(y | x) \leq E(y | x, z = 1)P(z = 1 | x ) + (1 - P(z = 1 | x)).
\] 

More generically, if we denote the minimum value the conditional 
outcome can take on as $\gamma_0$ and the maximum by $\gamma_1$, then 
the conditional mean can be bounded by 

\[
E(y | x, z = 1)P(z = 1 | x) + \gamma_0P(z=0|x) \leq E(y | x) \leq E(y | x, z = 1)P(z = 1 | x ) + \gamma_1 P(z = 0 | x).
\]

The conditional expectations in these expressions can be estimated 
parametrically or non-parametrically, and confidence intervals can be 
derived for the locations of the endpoints.

## Example

Using simulated data we present a simple example of the `pidoutcomes()` function
which we use to estimate partially identified conditional conditional means.

```{r example, cache=FALSE}
## We generate some fake data
library(rbounds)
set.seed(42)

N <- 100
x <- rnorm(N)
z <- rbinom(N, 1, pnorm(x))
y <- rbinom(N, 1, pnorm(x))

df <- data.frame(y, x, z)

res <- pidoutcomes(y~x, z, df)
res

# We can plot it with a generic plotting method
plot(res)

```

## Bounding Average Treatment Effects

The `ate_bounds()` function computes Manski worst-case bounds on the
Average Treatment Effect (ATE) when treatment assignment may be non-random.

The ATE is $\tau = E[Y(1)] - E[Y(0)]$ where $Y(1)$ and $Y(0)$ are
potential outcomes. Since we only observe $Y(1)$ for treated units and
$Y(0)$ for control units, the ATE is not point-identified without assumptions.

If outcomes are bounded (e.g., $Y \in [0,1]$), we can derive worst-case bounds:

$$\tau^L = E^L[Y(1)] - E^U[Y(0)]$$
$$\tau^U = E^U[Y(1)] - E^L[Y(0)]$$

### ATE Example

```{r ate_example, cache=FALSE}
# Simulate data with selection into treatment
set.seed(123)
N <- 200
x <- rnorm(N)

# Treatment depends on x (selection)
d <- rbinom(N, 1, pnorm(0.5 * x))

# Potential outcomes
y0 <- rbinom(N, 1, pnorm(-0.3 + 0.3 * x))
y1 <- rbinom(N, 1, pnorm(0.3 + 0.3 * x))

# Observed outcome
y <- d * y1 + (1 - d) * y0
df <- data.frame(y = y, x = x, d = d)

# True ATE (in this simulation we can compute it)
true_ate <- mean(y1 - y0)
cat("True ATE:", round(true_ate, 3), "\n")

# Estimate ATE bounds
ate_res <- ate_bounds(y ~ x, treatment = d, data = df)
ate_res

# Plot CATE bounds
plot(ate_res)
```
