---
output:
  md_document:
    variant: markdown_github
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# rbounds: Estimating Bounds on Conditional Expectation Functions

The goal of rbounds is to estimate bounds on the location of parameters
in models in which the value of the parameter can only be partially
identified. 

The package was developed to estimate the conditional mean of a 
binary outcome with non-random missing outcomes.
At the moment the package contains one function which estimates
bounds on a conditional mean using non-parametric kernel regression
estimators from the `np` package. In the future I hope to extend the
package to cover treatment effect estimation, and to work with some
additional assumptions such as monotone regresion and monotone
treatment response.

## Example of bounding the conditional mean

Imagine that we are interested in estimating
$E(y | x)$ where $y \in (0,1)$. Unfortunately, we are 
missing some observations of $y$. Whether an observation
is missing is denoted by a dummy variable $z$, where when $z=1$
$y$ is observed and otherwise the value of $y$ is not observed.
We have some covariates $x$ whose value we observe for all observations.

By the law of total probability it follows that 

\[E(y | x) = E(y | x, z = 1)P(z = 1 | x) + E(y | x, z = 0)P(z = 0 | x).\]

Each of the pieces of this equation can be estimated from 
the available data with the exception of $E(y|x, z=0)$, which
is the conditional expectation of the outcome for the cases in 
which the outcome value is unobserved. 

However, since $0 \leq y \leq 1$, the unobserved
function can take on the value of at most $1$ -- if all the
unobserved outcomes were 1, or at least $0$. Thus the true
conditional mean must fall within the bounds:

\[
E(y | x, z = 1)P(z = 1 | x) \leq E(y | x) \leq E(y | x, z = 1)P(z = 1 | x ) + (1 - P(z = 1 | x)).
\] 

More generically, if we denote the minimum value the conditional 
outcome can take on as $\gamma_0$ and the maximum by $\gamma_1$, then 
the conditional mean can be bounded by 

\[
E(y | x, z = 1)P(z = 1 | x) + \gamma_0P(z=0|x) \leq E(y | x) \leq E(y | x, z = 1)P(z = 1 | x ) + \gamma_1 P(z = 0 | x).
\]

The conditional expectations in these expressions can be estimated 
parametrically or non-parametrically, and confidence intervals can be 
derived for the locations of the endpoints.

## Example

Using simulated data we present a simple example of the `pidoutcomes()` function
which we use to estimate partially identified conditional conditional means.

```{r example}
## We generate some fake data
library(rbounds)
set.seed(42)
N <- 100
x <- rnorm(N)
z <- rbinom(N, 1, pnorm(x))
y <- rbinom(N, 1, pnorm(x))

df <- data.frame(y, x, z)

res <- pidoutcomes(y~x, z, df)
res

# We can plot it with a generic plotting method
plot(res)

```
